### 2024-02-02 thesis/abstract.tex
added thesis, (the first commit is uploading the folder of the master thesis preparation latex)

\begin{abstract}
    The theory of kernel methods will be applied to the problem of probabilistic forecasting taking as input some data.
    Conceptually, the proposed analysis could be applied to any kind of data. In this study we considered the electricity market, because of the interesting implications  on risk management tasks.

  %This example thesis briefly shows the main features of our thesis
  %style, and how to use it for your purposes.
\end{abstract}