### 2024-03-06 thesis/input_files/energymarket.tex
6 march

Chapter explaining the energy market

- what are smart grids?

- introduction of smart grids

- renewable integration requirements

\includegraphics[width=\textwidth]{images/electricity-production-by-source.png}

- How auctions work

- Difference intraday dayhead all other stuff

- EPEX Nordpool

- Prices can be negative