### 2024-03-06 thesis/input_files/symbols.tex

\glsxtrnewsymbol[description={acceleration}]{a}{\ensuremath{a}}

%C
\glsxtrnewsymbol[description={Continous ranked probability score}]{CRPS}{CRPS}

%E
\glsxtrnewsymbol[description={Exploratory data analysis}]{EDA}{EDA}
\glsxtrnewsymbol[description={Electric forecasting}]{EF}{EF}
\glsxtrnewsymbol[description={Electric vehicles}]{EVs}{EVs}
\glsxtrnewsymbol[description={Extract transform load}]{ETL}{ETL}

%L
\glsxtrnewsymbol[description={Low carbon technologies}]{LCTs}{LCTs}

\glsxtrnewsymbol[description={Low voltage}]{LV}{LV}

%P
\glsxtrnewsymbol[description={Prediction interval}]{PI}{PI}
\glsxtrnewsymbol[description={Price Forecasting}]{PF}{PF}

\glsxtrnewsymbol[description={Probabilistic Price Forecasting}]{PPF}{PPF}

%S
\glsxtrnewsymbol[description={Sustainable development goals}]{SDGs}{SDGs}

%
\glsxtrnewsymbol[description={Transmission system operator}]{TSO}{TSO}