### 2024-03-09 thesis/input_files/symbols.tex
9 march 2024

%For mathematical terms \ensuremath
% \glsxtrnewsymbol[description={Autoregressive exogenous model}]{ARX}{\ensuremath{ARX}}

%A
\glsxtrnewsymbol[description={Artificial neural network}]{ANN}{ANN}
\glsxtrnewsymbol[description={Autoregressive model}]{AR}{AR}
\glsxtrnewsymbol[description={Autoregressive moving average model}]{ARMA}{ARMA}
\glsxtrnewsymbol[description={Autoregressive integrated moving average model}]{ARIMA}{ARIMA}
\glsxtrnewsymbol[description={Autoregressive exogenous model}]{ARX}{ARX}

%C
\glsxtrnewsymbol[description={Computational intelligence}]{CI}{CI}
\glsxtrnewsymbol[description={Conditional kernel density}]{CKD}{CKD}
\glsxtrnewsymbol[description={Continous ranked probability score}]{CRPS}{CRPS}

%D
\glsxtrnewsymbol[description={Distributional neural network}]{DDNN}{DDNN}

%E
\glsxtrnewsymbol[description={Exploratory data analysis}]{EDA}{EDA}
\glsxtrnewsymbol[description={Electricity forecasting}]{EF}{EF}
\glsxtrnewsymbol[description={Electricity load forecasting}]{ELF}{ELF}
\glsxtrnewsymbol[description={Empirical mode decomposition}]{EMD}{EMD}
\glsxtrnewsymbol[description={Electricity load forecasting}]{EPF}{EPF}
\glsxtrnewsymbol[description={Exponential smoothing models}]{ESM}{ESM}
\glsxtrnewsymbol[description={Extract transform load}]{ETL}{ETL}
\glsxtrnewsymbol[description={Electric vehicles}]{EVs}{EVs}

%G
\glsxtrnewsymbol[description={Gradient boosting regression tree}]{GBRT}{GBRT}
\glsxtrnewsymbol[description={Global energy forecasting competition}]{GEFCom}{GEFCom}

%H
\glsxtrnewsymbol[description={Holt-Winters-Taylor exponential smoothing method}]{HWT}{HWT}

%K
\glsxtrnewsymbol[description={Kernel density estimation}]{KDE}{KDE}

%L
\glsxtrnewsymbol[description={Low carbon technologies}]{LCTs}{LCTs}
\glsxtrnewsymbol[description={Least squares support vector regression}]{LSSVR}{LSSVR}

\glsxtrnewsymbol[description={Low voltage}]{LV}{LV}

%M
\glsxtrnewsymbol[description={Multiple linear regression}]{MLR}{MLR}

%O
\glsxtrnewsymbol[description={Ordinary least squares}]{OLS}{OLS}

%R
\glsxtrnewsymbol[description={Random forest}]{RF}{RF}
\glsxtrnewsymbol[description={Reproducing kernel hilbert space}]{RKHS}{RKHS}

%P
\glsxtrnewsymbol[description={Prediction interval}]{PI}{PI}
\glsxtrnewsymbol[description={Price Forecasting}]{PF}{PF}

\glsxtrnewsymbol[description={Probabilistic Price Forecasting}]{PPF}{PPF}

%Q
\glsxtrnewsymbol[description={Quantile regression}]{QR}{QR}

\glsxtrnewsymbol[description={Quantile regression averaging}]{QRA}{QRA}

%S
\glsxtrnewsymbol[description={Small and medium-sized enterprises}]{SME}{SME}

\glsxtrnewsymbol[description={Smoothed nonparametric ARX}]{SNARX}{SNARX}

\glsxtrnewsymbol[description={Sustainable development goals}]{SDGs}{SDGs}
\glsxtrnewsymbol[description={Support vector machine}]{SVM}{SVM}

%T
\glsxtrnewsymbol[description={Time of use tariffs}]{TOU}{TOU}
\glsxtrnewsymbol[description={Transmission system operator}]{TSO}{TSO}