### 2024-03-10 thesis/abstract.tex
March 10th 2024

% PRELIMINARY ABSTRACT, ONCE I AM DONE WRITE THE FINAL ABSTRACT
\begin{abstract}
    The theory of kernel methods will be applied to the problem of point and probabilistic forecasting for the energy sector.
    Such choice is motivated by its interesting implications.

    % on risk management tasks.

    % - Contribution: Kernel herding applied to the
    % field of probabilistic forecasting

  %This example thesis briefly shows the main features of our thesis
  %style, and how to use it for your purposes.
\end{abstract}