### 2024-03-20 thesis/input_files/symbols.tex
March 20th 2024

%For mathematical terms \ensuremath
% \glsxtrnewsymbol[description={Autoregressive exogenous model}]{ARX}{\ensuremath{ARX}}

%A
\glsxtrnewsymbol[description={Artificial neural network}]{ANN}{ANN}
\glsxtrnewsymbol[description={Autoregressive model}]{AR}{AR}
\glsxtrnewsymbol[description={Autoregressive moving average model}]{ARMA}{ARMA}
\glsxtrnewsymbol[description={Autoregressive integrated moving average model}]{ARIMA}{ARIMA}
\glsxtrnewsymbol[description={Autoregressive exogenous model}]{ARX}{ARX}

%C
\glsxtrnewsymbol[description={Computational intelligence}]{CI}{CI}
\glsxtrnewsymbol[description={Conditional kernel density}]{CKD}{CKD}
\glsxtrnewsymbol[description={Continous ranked probability score}]{CRPS}{CRPS}

%D
\glsxtrnewsymbol[description={Distributional neural network}]{DDNN}{DDNN}

%E
\glsxtrnewsymbol[description={Exploratory data analysis}]{EDA}{EDA}
\glsxtrnewsymbol[description={Electricity forecasting}]{EF}{EF}
\glsxtrnewsymbol[description={Electricity load forecasting}]{ELF}{ELF}
\glsxtrnewsymbol[description={Empirical mode decomposition}]{EMD}{EMD}
\glsxtrnewsymbol[description={Electricity load forecasting}]{EPF}{EPF}
\glsxtrnewsymbol[description={Exponential smoothing models}]{ESM}{ESM}
\glsxtrnewsymbol[description={Extract transform load}]{ETL}{ETL}
\glsxtrnewsymbol[description={Electric vehicles}]{EVs}{EVs}

%G
\glsxtrnewsymbol[description={Gradient boosting regression tree}]{GBRT}{GBRT}
\glsxtrnewsymbol[description={Global energy forecasting competition}]{GEFCom}{GEFCom}
\glsxtrnewsymbol[description={Gaussian process regression}]{GPR}{GPR}

%H
\glsxtrnewsymbol[description={Holt-Winters-Taylor exponential smoothing method}]{HWT}{HWT}

%K
\glsxtrnewsymbol[description={Kernel density estimation}]{KDE}{KDE}

%L
\glsxtrnewsymbol[description={Low carbon technologies}]{LCTs}{LCTs}
\glsxtrnewsymbol[description={Least squares support vector regression}]{LSSVR}{LSSVR}

\glsxtrnewsymbol[description={Low voltage}]{LV}{LV}

%M
\glsxtrnewsymbol[description={Mean absolute error}]{MAE}{MAE}
\glsxtrnewsymbol[description={Mean absolute scaled error}]{MASE}{MASE}
\glsxtrnewsymbol[description={Mean absolute percentage error}]{MAPE}{MAPE}
\glsxtrnewsymbol[description={Minimal gated memory network}]{MGM}{MGM}

\glsxtrnewsymbol[description={Multiple linear regression}]{MLR}{MLR}

%O
\glsxtrnewsymbol[description={Ordinary least squares}]{OLS}{OLS}

%R
\glsxtrnewsymbol[description={Random forest}]{RF}{RF}
\glsxtrnewsymbol[description={Reproducing kernel hilbert space}]{RKHS}{RKHS}
\glsxtrnewsymbol[description={Root mean squared error}]{RMSE}{RMSE}

%P
\glsxtrnewsymbol[description={Prediction interval}]{PI}{PI}
\glsxtrnewsymbol[description={Price Forecasting}]{PF}{PF}

\glsxtrnewsymbol[description={Probabilistic Price Forecasting}]{PPF}{PPF}

%Q
\glsxtrnewsymbol[description={Quantile regression}]{QR}{QR}

\glsxtrnewsymbol[description={Quantile regression averaging}]{QRA}{QRA}

%S
\glsxtrnewsymbol[description={Small and medium-sized enterprises}]{SME}{SME}

\glsxtrnewsymbol[description={Smoothed nonparametric ARX}]{SNARX}{SNARX}

\glsxtrnewsymbol[description={Sustainable development goals}]{SDGs}{SDGs}
\glsxtrnewsymbol[description={Support vector machine}]{SVM}{SVM}

%T
\glsxtrnewsymbol[description={Time of use tariffs}]{TOU}{TOU}
\glsxtrnewsymbol[description={Transmission system operator}]{TSO}{TSO}

%MATHEMATICAL NOTATION
% D
\glsxtrnewsymbol[description={dual optimum}]{d^*}{\ensuremath{d^*}}

% F
\glsxtrnewsymbol[description={feature matrix}]{Phi(x)}{\ensuremath{\Phi(x)}}

\glsxtrnewsymbol[description={feature space}]{F}{\ensuremath{F}}

\glsxtrnewsymbol[description={feature vector}]{phi(x)}{\ensuremath{\phi(x)}}

% G
\glsxtrnewsymbol[description={dual function}]{g(v)}{\ensuremath{g(v)}}

% K
\glsxtrnewsymbol[description={kernel matrix}]{K}{\ensuremath{K}}

\glsxtrnewsymbol[description={kernel function}]{k(x,x')}{\ensuremath{k(x,x')}}

% L
\glsxtrnewsymbol[description={lagrangian}]{L}{\ensuremath{L}}

% P
\glsxtrnewsymbol[description={pinball loss}]{rho_q}{\ensuremath{\rho_q}}

\glsxtrnewsymbol[description={primal optimum}]{p^*}{\ensuremath{p^*}}

% R
\glsxtrnewsymbol[description={relative interior}]{relint}{\textrm{relint}}

\glsxtrnewsymbol[description={reproducing kernel hilbert space}]{H}{\ensuremath{H}}