### 2024-03-22 thesis/input_files/energymarket.tex
decision trees

Chapter explaining the energy market
Chapter 9 of https://pubs.naruc.org/pub.cfm?id=536E10A7-2354-D714-5191-A8AAFE45D626

- what are smart grids?

- introduction of smart grids

- renewable integration requirements

\includegraphics[width=\textwidth]{images/electricity-production-by-source.png}

- How auctions work

- Difference intraday dayhead all other stuff

- EPEX Nordpool

- Prices can be negative