### 2024-04-29 thesis/input_files/paper1.tex
comm 3.1

% Some commands used in this file

\section{Kernel mean embedding of distributions: A review and beyond}
% \href{https://arxiv.org/abs/1605.09522}{
From this first paper \cite{Muandet_2017}, the notation and terms used in the theory of Reproducing Kernel Hilbert Spaces are summarized.
\\
Many algorithms use the inner product as similarity measure between data instances $x, x' \in \mathcal{X}$. However, this inner product spans only the class of linear similarity measures. 
\\
The idea behind kernel methods is to apply a non-linear transformation $\phi$ to the data $x$ 
in order to get a more powerful non linear similarity measure.
\begin{align*}
\phi(x)\colon \mathcal{X} &\to \mathcal{F}
    \\
    x&\to \phi(x)
\end{align*}

We take the inner product in the high dimensional space $\mathcal{F}$ mapped by $\phi(x)$, i.e.\
\\
\begin{align*}
k(x,x'):=&\langle \phi(x), \phi(x') \rangle_{\mathcal{F}}
\end{align*}
\\
where $\phi(x)$ is referred to as feature map while $k$ is the kernel function.
\\
Therefore, we can kernelize any algorithm involving a dot product by substituting $\langle x, x' \rangle_{\mathcal{X}}$ with $\langle \phi(x), \phi(x') \rangle_{\mathcal{F}}$
\\
One would expect constructing the feature maps explicitly and then evaluating their inner product in $\mathcal{F}$ to be computationally expensive, and indeed it is. However, we do not have to explicitly perform such calculations. This is because of the kernel trick.
To illustrate the idea behind the kernel trick consider the following example. 
\\
Suppose $x \in \mathbb{R}^2$ and assume $\phi(x)=(x_{1}^{2},x_{2}^{2},\sqrt{2}x_{1}x_{2})$, then the inner product in the feature space is $x_{1}^{2}x_{1}^{'2},x_{2}^{2}x_{2}^{'2}+2x_{1}x_{2}x'_{1}x'_{2}$.
Notice that this is the same of $\langle \phi(x), \phi(x') \rangle$; thus the kernel trick consists of just using $k(x,x')=:(x^Tx')^2$.

\subsection{Reproducing kernel hilbert space}
Subsequently we introduce the definitions that make up the basis for the theory of kernel methods.

\begin{definition}
    A sequence $\{v_n\}_{n=1}^{\infty}$ of elements of a normed space $\mathcal{V}$ is a Cauchy sequence if for every $\epsilon>0$, there exist $N=N(\epsilon) \in \mathbb{N}$ such that $\|v_n-v_m\|_{\mathcal{V}}<\epsilon \ \ \forall m,n\geq N$  
\end{definition}

\begin{definition}
    A complete metric space is a metric space in which every Cauchy sequence is convergent.
\end{definition}

\begin{definition}
    A Hilbert space is a vector space $\mathcal{H}$ with an inner product $\langle \cdot, \cdot \rangle$ such that the norm defined by $\|f\|=\sqrt{\langle f, f \rangle}$
turns $\mathcal{H}$ into a complete metric space.
\end{definition}

\begin{definition}
    Let $(\mathcal{H}, \langle \cdot, \cdot \rangle_\mathcal{H})$ be a Hilbert space of real-valued functions on $\mathcal{X}$. A function $k: \mathcal{X} \times \mathcal{X} \to \mathbb{R}$ is called a reproducing kernel of $\mathcal{H}$ if and only if 
     \begin{align}
        k(x, \cdot) \in \mathcal{H} &\quad \text{for all   } x \in \mathcal{X}    \\
        \langle  h, k(x, \cdot) \rangle_\mathcal{H} = h(x) &\quad \text{for all   } h \in \mathcal{H}, x \in \mathcal{X}
    \end{align}
\end{definition}

% \begin{definition}
%     RKHS.
%     A Reproducing Kernel Hilbert Space is a Hilbert space with the evaluation functionals $\mathcal{F}_{x}(f):=f(x)$ bounded, i.e. $\forall x \in \mathcal{X}$ there exists some $C>0$ such that $\| \mathcal{F}_{x}(f)\|=\|f(x)\| \leq C \|f\|_{\mathcal{H}} \ \forall f \in \mathcal{H}$
% \end{definition}

\begin{theorem}
    [Riesz Representation]. If $A : \mathcal{H} \rightarrow \mathbb{R}$ is a bounded linear operator on a Hilbert space $\mathcal{H}$ , there exists some $g_{A} \in \mathcal{H}$ such that $A(f) = \langle f,g_A\rangle_\mathcal{H}, \ \forall f \in \mathcal{H}$.
\end{theorem}

The Riesz representation theorem results in the following proposition for reproducing kernel hilbert spaces.
\begin{proposition}
For each $x \in \mathcal{X}$ there exists a function $k_{x} \in \mathcal{H}$ such that $\mathcal{F}_{x}(f)=\langle k_{x}, f\rangle_{\mathcal{H}}=f(x)$    
\end{proposition}

The function $k_{x}$ is the reproducing kernel evaluated at point $x$.
Furthermore, note that $k_{x}$ is itself a function lying in it.

\begin{align*}
    k_{x}(y)=\mathcal{F}_{y}(k_{x})=\langle k_{x}, k_{y}\rangle_{\mathcal{H}=\langle \phi(x), \phi(y)\rangle_{\mathcal{H}}}
\end{align*}

\subsection{Kernel families}
Table \ref{tab:kernel types} contains popular kernel families in literature and applications.
\begin{table}
    \caption{Kernel types}
    \begin{tabular}{lll}
        \toprule
       Kernel function & Equation & Hyperparameters \\
       \midrule
       Linear &  $k(x_1,x_2)=x_1x_2$ &   \\
       Polynomial &  $k(x_1,x_2)=(x_1^\intercal x_2+c)^d$ &   c, d\\
       Gaussian RBF &  $k(x_1,x_2)=e^{-\frac{\|x_1-x_2|\|^2}{2\sigma^2}}$ &   $\sigma$\\
       Exponential RBF/Laplacian& $k(x_1,x_2)=e^{-\frac{|x_1-x_2|}{\gamma}}$ &   $\gamma$\\
       Hyperbolic/Sigmoid Kernel &  $k(x_1,x_2)=\tanh(\gamma x_1^\intercal x_2+r)$ &  $\gamma$, r \\
       Periodic &  $k(x_1,x_2)=e^{\frac{-2 sin^2\left(\frac{\pi}{p}|x_1-x_2| \right)}{l^2}} $ &   p, l\\
       Chi-squared kernel &  $k(x_1,x_2)=e^{-\gamma \sum\limits_i \frac{(x_i - y_i)^2}{x_i+y_i}} $ &   $\gamma$\\
        Cosine  &  $k(x_1,x_2)=\frac{xy^\intercal}{\|x\|_{L_2}\|y\|_{L_2}}$ &   \\
       Matern  &  $k(x_1, x_2) =  \frac{1}{\gamma(\nu)2^{\nu-1}}\Bigg(
        \frac{\sqrt{2\nu}}{l} d(x_1 , x_2 )
        \Bigg)^\nu K_\nu\Bigg(
        \frac{\sqrt{2\nu}}{l} d(x_1 , x_2 )\Bigg)$ &  $\gamma, \nu$ \\

        \bottomrule
    \end{tabular}
    \label{tab:kernel types}
\end{table}